---
title: "week_4_methods2_exercise.Rmd"
author: "Sigurd Sørensen"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Math notation
While cognitive science is not a math degree several cog-sci papers and teaching books makes heavyli use of math notation. It is therefore important to get a fundamental understanding of math operators and notation so you will be able to read and understand formulas when they're presented to you. 

So to practise the reading of math notation I'll present you with a few lines and it is your job to interpret them and translate into plain english. We will also do it in the reverse order where I present you a statement and you've to translate it into math notation. 


Some of these examples are stupid but it is to get you comfy with the notation signs. 
## Exercises on notation 
If we wanna write math notation in markdowns we can make use of something called latex. Latex is a language for math notation that can be used in many editors and IDE. 

In R you can start a latex notation by using $. In the same way that you can start a R chunk with 

```{r setup, include=FALSE}

```

dollar signs initatite a latex chunk. 

You can use double dollar signs
$$
\sum a_i = \sum b _i
$$
or single
$ a = b $

It only changes how the code is presented in the markdown. 

### Lets get started with some exercises. 

$$  \sum x_i >> \sum y_i,   $$
__answer:__
The sum of the indexed variable,x, is much greater than the sum of indexed variable, y.

$$ X \in \mathbb{R}^+ $$
__answer:__
the variable X belongs to the group of positive real numbers

$$ Y \in \mathbb{I}^+ $$
__answer:__
the variable Y is in the group of positive imaginary numbers

$$ x \in \mathbb{R}^+ \therefore x \notin \mathbb{R}^- $$
__answer:__
variable,x, belongs to the positive real numbers, therefore x does not belong to the negative real numbers

$$ y > x \therefore x - y \in \mathbb{R}^-$$
__answer:__
y is greater than x, therefore y subtracted from x belongs to negative real numbers

$$ x \in \mathbb{I}^-  y \in \mathbb{I}^+ \therefore x + y  \in \mathbb{I} $$

__answer:__ 
x is a negative imaginary number and y is a postive negative number, therefore they are both in the integers numbers

$$ x \varpropto y \because x = c \cdot y  $$
__answer:__ 
x is proportional on y because x is the same as y times a proportional constant.

$$x \in \mathbb{I} \ and  \ x \notin \mathbb{I}^- \Rightarrow x \in \mathbb{I}^+ $$ 
__answer:__ 
x element of to integers number and x does not to be found in negative imaginary, concluding, x belongs to the positive imaginary numbers.


### Part 2
Now i'll present you for some statements that you should translate into logical statements using the notation learned from the chapter. 

For a overview of latex commands see https://www.caam.rice.edu/~heinken/latex/symbols.pdf 
and https://oeis.org/wiki/List_of_LaTeX_mathematical_symbols


##### lets go
Let x be the any number of positive integers and y be any negative integer. Therefore the product 
of x and y must be a negative integer. 
__answer:__
$$ x \in \mathbb{I}^+ \ and \ y \in \mathbb{I}^- \therefore x*y 	\in \mathbb{I}^-$$

x is equal to 0 and a exist in the positive real numbers therefore a to the power of x is equal to 1. 
__answer:__
$$ x=0 \  	\exists \  \mathbb{R}^+ \ \therefore \ a^x=1$$

The mean of y_hat is equal to the sum of of all y_i divided by the length of y_i which is an approximate of the mean of y.   
__answer:__
$$ \widehat{\bar{y}}= \sum{\frac{y_i}{|y_1|}}  \approx \overline{y}  $$

x is proportional to y iff (if and only if) x = y * c
__answer:__
$$ x \propto y \iff x=y*c$$

**Make up your own example.**
Make up a few of your own examples where you generate both the plain English and the math notation. 


We will walk through some of these examples together. Continue with exercises from the book. 
# Nested functions

# Sets
let x be a set {0,2,3,10,8,4,11}
let y be a set {1,3,10,4,20,11}

Find the following set z for each expression. 

  1)
$$ z = x \cup y  $$

  2)
$$ z =x \cap y $$

  3)
$$z = x \setminus y $$


  4) 
$$ max(x) $$



  5)
$$ max(x \cap y) $$


# Exercises from the book. 
These exercises is to refresh some math fundamentals from high-school which will be useful for coming weeks. You can do most of these with pen and paper. 


### 1.1 
Simplify the following expressions as much as possible:
(−x4y2)2 9(30) (2a2)(4a4)
x4
x3 (−2)7−4

1
27b3
1/3
y7y6y5y4 2a/7b
11b/5a
(z2)4

### 1.4 
The relationship between Fahrenheit and Centigrade can be expressed
as 5f − 9c = 160. Show that this is a linear function by putting it in
y = mx + b format with c = y. Graph the line indicating slope and
intercept.

### 1.6

Solve the following inequalities so that the variable is the only term
on the left-hand side:
x − 3 < 2x+ 15
11 − 4/3t > 3 
5/6y + 3(y − 1) ≤ 11/6 (1 − y) + 2y

### 1.9
 ! [Alt text](/Skærmbillede 2022-02-23 140005.png)
 


### 1.18 (hard)

### 1.21

### 1.22


# After you're done. 
If you finish look over all the other exercises and see if there is any that would trouble you. 
If so then try and work through it yourself and raise your hand for help. :) 