---
title: "Methods 2 -- Portfolio Assignment 1"
output:
  pdf_document: default
  html_notebook: default
---

- _Type:_ Group assignment
- _Due:_ 13 March 2022, 23:59

---

Please submit your answers on GitHub Classroom. The best way to give your answers is in a LaTeX or R Markdown file, but you can also create a PDF in some other way.

---

1) Given the vectors $\vec{u} = (1,1,1)$, $\vec{v} = (2,3,1)$, and $\vec{w} = (-1,-1,2)$, compute the following products:
    a) $\vec{u} \cdot \vec{v}$
    b) $\vec{u} \cdot \vec{w}$
    c) $\vec{v} \cdot \vec{w}$
    d) $\vec{u} \times \vec{v}$
    e) $\vec{u} \times \vec{w}$
    f) $\vec{v} \times \vec{w}$

```{r}
pacman::p_load(pracma)
u <- c(1,1,1)
v <- c(2,3,1)
w <- c(-1,-1,2)

#a)
u%*%v
#6

#b)
u%*%w
#0

#c)
v%*%w
#-3

#d)
cross(u,v)
#-2 1 1

#e)
cross(u,w)
# 3 -3 0

#f)
cross(v,w)
#7 -5 1
```


---

2) Vectors are _orthogonal_ (or _perpendicular_) when their dot product is zero: $\vec{x} \perp \vec{y} \Longleftrightarrow \vec{x} \cdot \vec{y} = 0$. A _unit vector_ is a vector with norm 1: $\Vert \vec{x} \Vert = 1$.
    a) Find a unit vector that is perpendicular to both $\vec{u} = (1,0,1)$ and $\vec{v} = (1,2,0)$.
    b) Find a vector that is orthogonal both to $\vec{u}_1 = (1,0,1)$ and $\vec{u}_2 = (1,3,0)$, and whose dot product with the vector $\vec{v} = (1,1,0)$ is equal to 8.
    
```{r}
#a)
x <- c(1,0,1)
y <- c(1,2,0)
x%*%y
cross(x,y) #finding a possible orothaganol vector
d <- c(-2,1,2)
d%*%y #checking dot produkt
x%*%d #checking dot product
#d is orthoganol on both vectors, but has a length of 3
#therefor the vector is -2/3,1/3,2/3


#b) 
a <- c(1,0,1)
b <- c(1,3,0)
c <- c(1,1,0)
cross(a,b)
d <- c(-3,1,3) #vector
a%*%d #orthoganal check
b%*%d #orthoganol check
c%*%d #result -2
d2 <- d*-4 #The final vector d(12,-4,-12)
c%*%d2 #result 8
```

---

3) Prove the geometric formula for the dot product $\vec{x} \cdot \vec{y} = \Vert \vec{x} \Vert \Vert \vec{y} \Vert \cos \varphi$, where $\varphi$ is the angle between the vectors $\vec{x}$ and $\vec{y}$.

Using the laws of Cosines, the properties of the dotproduct the following relation between $\vec{x}$ and $\vec{y}$ with an angle of $\varphi$ is proved:
$$
\vec{x} \cdot \vec{y} = \Vert \vec{x} \Vert \Vert \vec{y} \Vert \cos \varphi
\\
Dotproduct\ follows:\ \vec{a}\cdot\vec{a}=\Vert\vec{a}\Vert
\\
Cosinerelation: a^2=b^2 +c^2-2*bc\cos\varphi
\\
$$
The connecting vector has the length of the difference between the norm of two vectors $ \Vert\vec{x}\Vert$ and  $\Vert\vec{y}\Vert$ is $\Vert\vec{x}-\vec{y}\Vert$. 
$$
\\
\therefore (\Vert\vec{x}-\vec{y}\Vert)^2=\Vert\vec{x}\Vert^2  +\Vert\vec{y}\Vert^2-2*\Vert\vec{x}\Vert\Vert\vec{y}\Vert\cos\varphi \ \ (1)
\\
\implies (\Vert \vec{x}-\vec{y}\Vert)^2=(\vec{x}-\vec{y})(\vec{x}-\vec{y})
\\
=\vec{x} \cdot \vec{x} + \vec{y} \cdot \vec{y} -2 (\vec{x}\cdot \vec{y})
\\
=\Vert \vec{x}\Vert^2+\Vert\vec{y}\Vert^2 -2 (\vec{x} \vec{y}) \ \ (2)
\\
$$
Using  (2)  in (1) we get:
$$
\\
\Vert \vec{x}\Vert^2+\Vert\vec{y}\Vert^2 -2 (\vec{x} \vec{y}) =\Vert\vec{x}\Vert^2 +\Vert\vec{y}\Vert^2-2*\Vert\vec{x}\Vert\Vert\vec{y}\Vert\cos\varphi
\\
\implies -2 (\vec{x} \vec{y})=-2*\Vert\vec{x}\Vert\Vert\vec{y}\Vert\cos\varphi
\\
\implies \vec{x} \vec{y}=\Vert\vec{x}\Vert\Vert\vec{y}\Vert\cos\varphi
\\
\implies \vec{x} \cdot\vec{y}=\Vert\vec{x}\Vert\Vert\vec{y}\Vert\cos\varphi \blacksquare
\\

$$
---

4) For the matrix
$$
X = \begin{bmatrix}
0 & 0 & 1 \\
0 & 1 & 0 \\
1 & 0 & 1
\end{bmatrix},
$$
calculate $X^n$ for $n = 2, 3,4, 5$. Write a rule for calculating higher values of $n$.

```{r}
X <- 
```

