# Class 01
### Setup chunk
```{r}
setwd('.')
pacman::p_load(ggplot2,tidyverse)
```

## Exercises from ROS page 32
### 1.2 
**Sketching a regression model and data**: Figure 1.1b shows data corresponding to the fitted line
*y = 46.3+3.0x* with residual standard deviation 3.9, and values of *x* ranging roughly from 0 to
4%.
#### (a)
Sketch hypothetical data with the same range of *x* but corresponding to the line y = 30 + 10x
with residual standard deviation 3.9.
```{r}
#linear regression: y=b0+b1*x+error
#simulate x
x1 <- rnorm(1000,2,0.6)
hist(x1)

#simulate error
e1 <- rnorm(1000,0,3.9) #error trending around 0 with a sd of 3.9

#formular
y1 <- 30+10*x1+e1
plot(y1)
```

#### (b)
Sketch hypothetical data with the same range of *x* but corresponding to the line y = 30 + 10x
with residual standard deviation 10.
```{r}
#linear regression: y=b0+b1*x+error
#simulate x
x1 <- rnorm(1000,2,0.6) #same x
hist(x1)

#simulate error
e2 <- rnorm(1000,0,10)

#formular
y2 <- 30+10*x1+e2
plot(y2)
```

## Applying learned skills on your selected dataset:  (IDEA)
```{r prep data}
df <- read.csv('fetal_health.csv')
view(df)
ls(df)
df <-  na.omit(df)
```
 [1] discrete 15-87           = "abnormal_short_term_variability"                       
 [2] continuous 0-0.0019      = "accelerations"                                         
 [3] discrete                 = "baseline.value"                                       
 [4] discrete 1-3             = "fetal_health"                                          
 [5] continuous 0-0.47..      = "fetal_movement"                                        
 [6]                          = "histogram_max"                                         
 [7] discrete 73-182          = "histogram_mean"                                        
 [8] discrete 76-186          = "histogram_median"                                      
 [9]                          = "histogram_min"                                         
[10]                          = "histogram_mode"                                        
[11]                          = "histogram_number_of_peaks"                             
[12]                          =  "histogram_number_of_zeroes"                            
[13]                          = "histogram_tendency"                                    
[14]                          = "histogram_variance"                                    
[15] discrete 0-180           = "histogram_width"                                       
[16] continuous 0-0.015       = "light_decelerations"                                   
[17] continuous 0-50.7        = "mean_value_of_long_term_variability"                   
[18] continuous 2.0-7.0       = "mean_value_of_short_term_variability"                  
[19] continuouS 0-91          = "percentage_of_time_with_abnormal_long_term_variability"
[20] discrete 0-0.005         = "prolongued_decelerations"                              
[21] binary 0.001 or 0(0.000) = "severe_decelerations"                              
[22] continuous 0-0.015       = "uterine_contractions" 
### Use a descriptive statistics and visual investigation to give a preliminary look into the data. (Use some of the things from chapter 2) 
```{r}

```

### Create several models and discuss what each model reliably can infer anything about. Mixed effect, which random effects? Logistic regression? Interaction effects? Should we dummy code some variables?  
### Do model comparison.
### Test the model's prediction capability.
### Write up a short summary of your model.  

### 1.3 - *could be a thing to do* 
Goals of regression: Download some data on a topic of interest to you. Without graphing the data
or performing any statistical analysis, discuss how you might use these data to do the following
things:
##### (a)
Fit a regression to estimate a relationship of interest.
##### (b)
Use regression to adjust for differences between treatment and control groups.
##### (c)
Use a regression to make predictions.

### Do Exercise 2.3 in the book (Hard) hints. ( Summarise_at() , str_sub(), pivot_longer() )

### Upload the .Rmd to your GitHub. 